---
title: "A4_P1_Data2018"
author: "Annamaya Halskov-Jensen"
date: "13/11/2018"
output: html_document
---


```{r setup, include=FALSE}
getwd()
library(pacman)
p_load(tidyverse, ggplot2, groupdata2, crqa, gridExtra, gdata, lmerTest, lme4)
current = "/Users/annamayahalskov-jensen/Documents/Cognitive Science/3rd semester/Experimental Methods 3/Heart Rate"
```

## Analysing Heart Rate and Respiration data

The goal of this assignment is to first familiarize you with heart rate, and respiration data and their preprocessing. The second part explores how to analyze interpersonal coordination of these signals.

## Step by step suggestions to solve the assignment

### Exploring physiological signals

- Choose one pair (one pair, three conditions)
- Load the logs

```{r, loading data}

conv <- read.csv("Study3_G1_T3_Conversation.csv")
sync <- read.csv("Study3_G1_T1_Synchronous.csv")
turn <- read.csv("Study3_G1_T2_TurnTaking.csv")

conv <- rename.vars(conv, "TimeMs", "time")
sync <- rename.vars(sync, "TimeMs", "time")
turn <- rename.vars(turn, "TimeMs", "time")

```

- Produce a plot of the participants' respiration signal and a different one of the participants' HR signal (for inspecting whether the data is usable)
  N.B: remember the slides: artifacts, downsampling, scaling.
  N.B. The gridExtra::grid.arrange() function allows you to display the plots side by side. E.g. grid.arrange(plot1, plot2, plot3,  ncol=3)

PLOTS:
```{r, plotting}

### Respiration plots for all three conditions
resp_conv <- ggplot(conv, aes(time, Resp1)) + 
  geom_line(aes(time, Resp1), color = "#4D6C95") +
  geom_line(aes(time, Resp2), color = "#E99F94") +
  labs(x = "Time", 
       y = "Respiration", 
       title = "Respiration over time (Conversation)")

resp_sync <- ggplot(sync, aes(time, Resp1)) + 
  geom_line(aes(time, Resp1), color = "#4D6C95") +
  geom_line(aes(time, Resp2), color = "#E99F94") +
  labs(x = "Time", 
       y = "Respiration", 
       title = "Respiration over time (Synchronous)")

resp_turn <- ggplot(turn, aes(time, Resp1)) + 
  geom_line(aes(time, Resp1), color = "#4D6C95") +
  geom_line(aes(time, Resp2), color = "#E99F94") +
  labs(x = "Time", 
       y = "Respiration", 
       title = "Respiration over time (Turn Taking)")

### Heart reate plots for all three conditions
hr_conv <- ggplot(conv, aes(time, HR1)) + 
  geom_line(aes(time, HR1), color = "#4D6C95") +
  geom_line(aes(time, HR2), color = "#E99F94") +
  labs(x = "Time", 
       y = "Heart rate", 
       title = "Heart rate over time (Conversation)")

hr_sync <- ggplot(sync, aes(time, HR1)) + 
  geom_line(aes(time, HR1), color = "#4D6C95") +
  geom_line(aes(time, HR2), color = "#E99F94") +
  labs(x = "Time", 
       y = "Heart rate", 
       title = "Heart rate over time (Synchronous)")

hr_turn <- ggplot(turn, aes(time, HR1)) + 
  geom_line(aes(time, HR1), color = "#4D6C95") +
  geom_line(aes(time, HR2), color = "#E99F94") +
  labs(x = "Time", 
       y = "Heart rate", 
       title = "Heart rate over time (Turn Taking)")

resp_turn
resp_conv
resp_sync
hr_turn
hr_sync
hr_conv

```

DATA CLEANING:
```{r, downsampling}
### Downsampling to 2401 observations
conv_1 <- conv %>%
  group(n = 100, method = 'greedy') %>% 
  dplyr::summarise(time = mean(time,na.rm=T), 
                   HR1 = mean(HR1,na.rm=T), 
                   HR2 = mean(HR2,na.rm=T), 
                   Resp1 = mean(Resp1,na.rm=T), 
                   Resp2 = mean(Resp2,na.rm=T))

sync_1 <- sync %>%
  group(n = 100, method = 'greedy') %>% 
  dplyr::summarise(time = mean(time,na.rm=T), 
                   HR1 = mean(HR1,na.rm=T), 
                   HR2 = mean(HR2,na.rm=T), 
                   Resp1 = mean(Resp1,na.rm=T), 
                   Resp2 = mean(Resp2,na.rm=T))

turn_1 <- turn %>%
  group(n = 100, method = 'greedy') %>% 
  dplyr::summarise(time = mean(time,na.rm=T), 
                   HR1 = mean(HR1,na.rm=T), 
                   HR2 = mean(HR2,na.rm=T), 
                   Resp1 = mean(Resp1,na.rm=T), 
                   Resp2 = mean(Resp2,na.rm=T))

```


```{r, rescaling}
# Rescaling
conv_1$HR1 <- scale(conv_1$HR1)
conv_1$HR2 <- scale(conv_1$HR2)
conv_1$Resp1 <- scale(conv_1$Resp1)
conv_1$Resp2 <- scale(conv_1$Resp2)

sync_1$HR1 <- scale(sync_1$HR1)
sync_1$HR2 <- scale(sync_1$HR2)
sync_1$Resp1 <- scale(sync_1$Resp1)
sync_1$Resp2 <- scale(sync_1$Resp2)

turn_1$HR1 <- scale(turn_1$HR1)
turn_1$HR2 <- scale(turn_1$HR2)
turn_1$Resp1 <- scale(turn_1$Resp1)
turn_1$Resp2 <- scale(turn_1$Resp2)
```

MORE PLOTS:
```{r, respiration plots}
### Respiration plots for all three conditions
resp_conv_1 <- ggplot(conv_1, aes(time, Resp1)) + 
  geom_line(aes(time, Resp1), color = "#4D6C95") +
  geom_line(aes(time, Resp2), color = "#E99F94") +
  labs(x = "Time", 
       y = "Respiration", 
       title = "Respiration over time (Conversation)")

resp_sync_1 <- ggplot(sync_1, aes(time, Resp1)) + 
  geom_line(aes(time, Resp1), color = "#4D6C95") +
  geom_line(aes(time, Resp2), color = "#E99F94") +
  labs(x = "Time", 
       y = "Respiration", 
       title = "Respiration over time (Synchronous)")

resp_turn_1 <- ggplot(turn_1, aes(time, Resp1)) + 
  geom_line(aes(time, Resp1), color = "#4D6C95") +
  geom_line(aes(time, Resp2), color = "#E99F94") +
  labs(x = "Time", 
       y = "Respiration", 
       title = "Respiration over time (Turn Taking)")

resp_sync_1
resp_conv_1
resp_turn_1
```


```{r, heart rate plots}
### Heart reate plots for all three conditions
hr_conv_1 <- ggplot(conv_1, aes(time, HR1)) + 
  geom_line(aes(time, HR1), color = "#4D6C95") +
  geom_line(aes(time, HR2), color = "#E99F94") +
  labs(x = "Time", 
       y = "Heart rate", 
       title = "Heart rate over time (Conversation)")

hr_sync_1 <- ggplot(sync_1, aes(time, HR1)) + 
  geom_line(aes(time, HR1), color = "#4D6C95") +
  geom_line(aes(time, HR2), color = "#E99F94") +
  labs(x = "Time", 
       y = "Heart rate", 
       title = "Heart rate over time (Synchronous)")

hr_turn_1 <- ggplot(turn_1, aes(time, HR1)) + 
  geom_line(aes(time, HR1), color = "#4D6C95") +
  geom_line(aes(time, HR2), color = "#E99F94") +
  labs(x = "Time", 
       y = "Heart rate", 
       title = "Heart rate over time (Turn Taking)")

hr_conv_1
hr_turn_1
hr_sync_1

```

MORE DATA CLEANING:
```{r, function for removing artefacts}

#Creating a function for removing artefacts
cillitbang <- function(ts,threshold){
  ts[ts > (mean(ts,na.rm = T) +
             (threshold*sd(ts,na.rm = T))) |
       ts < (mean(ts,na.rm = T) -
             (threshold*sd(ts,na.rm = T)))] = mean(ts,na.rm = T)
  return(ts)
  }
```


```{r, removing artefacts for resp}
#Setting and applying threshold for respiration
threshold_resp = 2.5
conv_1$Resp1 <- cillitbang(conv_1$Resp1, threshold_resp)
conv_1$Resp2 <- cillitbang(conv_1$Resp2, threshold_resp)

sync_1$Resp1 <- cillitbang(sync_1$Resp1, threshold_resp)
sync_1$Resp2 <- cillitbang(sync_1$Resp2, threshold_resp)

turn_1$Resp1 <- cillitbang(turn_1$Resp1, threshold_resp)
turn_1$Resp2 <- cillitbang(turn_1$Resp2, threshold_resp)
```


```{r, removing artefacts for HR}
#Setting and applying a threshold for heart rate
threshold_hr = 2.75
conv_1$HR1 <- cillitbang(conv_1$HR1, threshold_hr)
conv_1$HR2 <- cillitbang(conv_1$HR2, threshold_hr)

sync_1$HR1 <- cillitbang(sync_1$HR1, threshold_hr)
sync_1$HR2 <- cillitbang(sync_1$HR2, threshold_hr)

turn_1$HR1 <- cillitbang(turn_1$HR1, threshold_hr)
turn_1$HR2 <- cillitbang(turn_1$HR2, threshold_hr)

```


- Can you eye-ball which condition if any displays more physiological coordination?
- Run crqa on heart rate and respiration data (find parameters, run crqa)

```{r, find parameters, run RQA}
#lgM = Maximum lag
#radius span = 2?     increasing radius step means smaller steps
#radius sample = 10?  number of radius points within the steps to be sampled

# Finding the optimal parameters:
par = list(lgM = 50, steps = seq(1, 6, 1), radiusspan = 100, radiussample = 40, normalize = 0, rescale = 0, mindiagline = 2, minvertline = 2, tw = 0, whiteline = F, recpt = F, fnnpercent = 10, typeami = "mindip")

#ans = optimizeParam(x1, x2, par, min.rec = 3.5, max.rec = 4.5)
par_conv_resp <- optimizeParam(na.omit(conv_1$Resp1), na.omit(conv_1$Resp2), par, min.rec = 2, max.rec = 5)
par_sync_resp <- optimizeParam(na.omit(sync_1$Resp1), na.omit(sync_1$Resp2), par, min.rec = 2, max.rec = 5)
par_turn_resp <- optimizeParam(na.omit(turn_1$Resp1), na.omit(turn_1$Resp2), par, min.rec = 2, max.rec = 5)

par_conv_hr <- optimizeParam(na.omit(conv_1$HR1), na.omit(conv_1$HR2), par, min.rec = 2, max.rec = 5)
par_sync_hr <- optimizeParam(na.omit(sync_1$HR1), na.omit(sync_1$HR2), par, min.rec = 2, max.rec = 5)
par_turn_hr <- optimizeParam(na.omit(turn_1$HR1), na.omit(turn_1$HR2), par, min.rec = 2, max.rec = 5)

# Choose a common value for delay, emb.dim and radius
results = crqa(g7_conv_1$Resp1, g7_conv_1$Resp2, delay = delay, embed = emddim, radius = radius, normalize = 0, rescale = 0, mindiagline = 2, minvertline = 2)
```


```{r, plotting RQA and stuff}
# Represent the plot:
RP = results$RP
RP = matrix(as.numeric(RP), nrow = ncol(RP))
cols = c("white","blue4")
image(RP, xlab = "", ylab = "", col = cols)

# Explore the lags of coordination:
Profile = drpdfromts(g7_conv_1$Resp1, g7_conv_1$Resp2, datatype = 'continuous', ws=50, radius=radius)
timecourse = round( seq(-5000,5000,100)/1000, digit = 1)
maxlag = Profile$maxlag/1000
profile = Profile$profile*100
Prof = data.frame(profile)

# Final plot
ggplot(Prof, aes(timecourse,profile))+geom_line()+ geom_vline(xintercept = timecourse[maxlag], colour='red')
```


- Does this tell you more than just eyeballing the plots?

### Systematically pre-process the data
- Loop through all the files (either with a loop or with a function), check which files should be excluded, if any, and save the pre-processed time-series. 
  Tip: plot and visually inspect the data to figure out which should be excluded.
  
```{r, collecting files}
##list of file names
files <- list.files(path = current, pattern="*.csv", full.names = F)
```


```{r, data-loading function, JanCo solution}
read_heart <- function(filename){
    raw <- read_csv(filename,
                    col_types = list(TimeMs = col_double(),
                                     Resp1 = col_double(),
                                     Resp2 = col_double(),
                                     ECG1 = col_double(),
                                     ECG2 = col_double(),
                                     HR1 = col_double(),
                                     HR2 = col_double()))
    name <- as.character(filename)
    
    data <- cbind(raw, name)
    
    data <- data %>%
      mutate(name = str_remove_all(name, "data/"),
         name = str_remove_all(name, ".csv")) %>%
      
      mutate(study = substr(name, 6, 6),
         group = substr(name, 9, 10),
         group = str_remove_all(group, "_"),
         trial = substr(name, 12, 12),
         condition = substr(name, 14, 30),
         condition = str_remove_all(condition, "_")) %>%
      
      select(-name)
    
    return(data)
}
```


```{r, loading all data, JanCo solution}
data_full_janco <- map_df(files, read_heart) %>%
   mutate(trial = factor(trial),
         condition = factor(condition),
         study = factor(study),
         group = factor(group))

data_full_janco <- rename.vars(data_full_janco, "TimeMs", "time")

#JanCo notes:
  #filter out 13820 NAs that appread exclusively in S2_G7_T3_TurnTaking
  #we're not discarding any infromation, there are 13820 NAs in every column
  #filter(!is.na(time)) 
```


```{r, test set, JanCo solution}
s1_g1_files <- as.data.frame(files) %>%
  filter(str_detect(files, "Study1")) %>%
  filter(str_detect(files, "G1"))

s1_g1_files <- as.character(s1_g1_files$files)

d <- map_df(s1_g1_files, read_heart) %>%
   mutate(trial = factor(trial),
         condition = factor(condition),
         study = factor(study),
         group = factor(group))
```


```{r, data cleaning, JanCo solution}
data_down <- data_full_janco %>%
  group(n = 100, method = 'greedy') %>% 
  dplyr::summarise(time = mean(time,na.rm=T), 
                   HR1 = mean(HR1,na.rm=T), 
                   HR2 = mean(HR2,na.rm=T), 
                   Resp1 = mean(Resp1,na.rm=T), 
                   Resp2 = mean(Resp2,na.rm=T))

data_down$HR1 <- scale(data_down$HR1)
data_down$HR2 <- scale(data_down$HR2)
data_down$Resp1 <- scale(data_down$Resp1)
data_down$Resp2 <- scale(data_down$Resp2)

data_down$HR1 <- cillitbang(data_down$HR1, threshold_hr)
data_down$HR2 <- cillitbang(data_down$HR2, threshold_hr)

data_down$Resp1 <- cillitbang(data_down$Resp1, threshold_resp)
data_down$Resp2 <- cillitbang(data_down$Resp2, threshold_resp)

```


- Run crqa on all the pre-processed time-series and save the output (don't forget to add columns with study, group, condition and trial). 

```{r}
#Function that downsamples, rescales and  removes outliers.
#Then print plots for HR and respiration in one 
#Lastly, creates variables with optimal parameters for crqa. 

#Output = dataframe and prints plots

vanish = function(filename, graphs = T, bangbycillit = T){
  
  #load file as csv
  df = read_csv(filename)
  df <- rename.vars(df, "TimeMs", "time")
  
  #Downsample
  df = df %>%    
  group(n= 100, method= 'greedy') %>%    
  dplyr::summarise(time= mean(time,na.rm=T), HR1 = mean(HR1,na.rm=T), HR2 = mean(HR2,na.rm=T), Resp1 =  mean(Resp1,na.rm=T), Resp2 = mean(Resp2,na.rm=T)) 
   
  #Rescale and create this in a new scaled variable
  df$Resp1=scale(df$Resp1)  
  df$Resp2=scale(df$Resp2)  
  df$HR1=scale(df$HR1) 
  df$HR2=scale(df$HR2)
  
  if(bangbycillit == T) {
  #Removing outliers using thresholds from before
  df$HR1=cillitbang(df$HR1,threshold_hr)
  df$HR2=cillitbang(df$HR2, threshold_hr)
  df$Resp1=cillitbang(df$Resp1, threshold_resp)
  df$Resp2=cillitbang(df$Resp2, threshold_resp)
  }
  
  
  #Adding columns with study identification
  df$study = str_extract(filename, "Study(\\d)") 
  df$group = str_extract(filename, "G(\\d+)")
  df$trial = str_extract(filename, "T(\\d)")
  
  #gsub  - {n}: matches exactly n times.
  # $: matches the end of the string.
  
  df$condition = gsub('.{4}$', '', strsplit(filename, "_")[[1]][4])
  df$filename = filename
  
  ##Calculating optimal parameters
    par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2, minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip")
  
  #Parameters for HR data
  
  optparamHR = try(optimizeParam(df$HR1, df$HR2, par, min.rec = 2, max.rec = 5))
  
  #Using loop from before, if the optimize function succeeded, save the results to the dataframe - if not, put NAs 
 
   if (length(optparamHR) > 1) {
   
  #need to unlist the parameters otherwise they can't be used for later calculations (no idea why it makes them a list in the first place)

   df$optRadiusHR = optparamHR$radius
   df$optEmbdimHR = optparamHR$emddim
   df$optDelayHR = optparamHR$delay 
   
   } else {
  
   df$optRadiusHR = NA
   df$optEmbdimHR = NA
   df$optDelayHR = NA
   }
  
  #Now for respiration data
  
  optparamResp = try(optimizeParam(df$Resp1, df$Resp2, par, min.rec = 2, max.rec = 5))
 
  #if the optimize function succeeded, save the results to the dataframe - if not, put NAs 
  if (length(optparamResp) > 1) {
    
    df$optRadiusResp = optparamResp$radius
    df$optEmbdimResp = optparamResp$emddim
    df$optDelayResp = optparamResp$delay

    } else {
    
    df$optRadiusResp = NA
    df$optEmbdimResp = NA
    df$optDelayResp = NA
    }
  
 
  #Writing new csv file with the new columns
  name = str_extract(filename, '.*(?=\\.csv)') #removing .csv from filename
  name = paste(name, "preprocessed.csv", sep = '_')
  
  #creating new directing to put files in 
  dir.create("Preprocessed", showWarnings = FALSE) #stops warnings if folder already exists
  
  #writing file to the new folder
  write.csv(df, file.path("Preprocessed", name), row.names=FALSE)
  if(graphs == T) {
  
  #Making plots 
  HR = ggplot(df, aes(time, HR1)) + 
    geom_line() + 
    geom_line(aes(time, HR2, color = "red")) + 
    ggtitle(filename) +
    theme(legend.position = "none")
  Resp = ggplot(df, aes(time, Resp1)) + 
    geom_line() + 
    geom_line(aes(time, Resp2, color = "red")) + 
    ggtitle(filename) +
  
      theme(legend.position = "none")
  
  #Set plots together
  grid.arrange(HR, Resp)
  
  #With arrangeGrob we can save to variable and then to disk
  plots = arrangeGrob(HR, Resp)
  
  #Giving each plot individual names for each plot based on the filename
  plotname = str_extract(filename, '.*(?=\\.csv)') #removing .csv from filename
  plotname = paste(plotname, "PLOT.png", sep = '_')
  
  #creating a folder for the plots 
  dir.create("Plots", showWarnings = FALSE) #stops warnings if folder already exists
  
  #writing file to the new folder
  
  #ggsave(plotname, plot = plots (the arrangeGrob variables), path = "Plots/")
  ggsave(file=plotname, plots, path = "Plots/")
  }
  
  #Return df
  return(df)
}


#Trying out the function (with jomfrurejsen)
#titanic = preproz("Study2_G5_T3_Conversation.csv")
#titanic = bind_rows(testout)

#Running all data in the beautiful function
all_preprocessed = lapply(files, vanish)
all_preprocessed = bind_rows(all_preprocessed)
#write_csv(all_preprocessed, "preprocessed_from_vanish.csv")

```


```{r, excluding troublesome data}
newdata <- all_preprocessed[!(all_preprocessed$condition=="Synchronous" & all_preprocessed$group=="G9"),]
#write.csv(newdata, "data_exclusive.csv")
```


```{r, getting median optimal parameters}
#Using median (more robust to outliers)
#NB: note that delay is not allowed to be 0, then set it to 1! 
optradius_resp <- median(newdata$optRadiusResp, na.rm = T)
optemddim_resp <- median(newdata$optEmbdimResp, na.rm = T)
optdelay_resp <- median(newdata$optDelayResp, na.rm = T)

#optradius_hr <- median(newdata$optRadiusHR, na.rm = T)
optradius_hr <- 2.65
optemddim_hr <- median(newdata$optEmbdimHR, na.rm = T)
optdelay_hr <- median(newdata$optDelayHR, na.rm = T)

```


```{r, crqa}
newdata <- newdata %>%
  mutate(group_trial = str_c(group, trial, sep = " "))

pairs <- unique(newdata$group_trial)

par = list(lgM = 50, steps = seq(1, 6, 1), radiusspan = 100, radiussample = 40, normalize = 0, rescale = 0, mindiagline = 2, minvertline = 2, tw = 0, whiteline = FALSE, recpt = FALSE, fnnpercent = 10, typeami = "mindip")

crqa_df_hr <- NULL
crqa_df_resp <- NULL

set.seed(13)

# Loop for HR
for (i in pairs){
  currentpair = filter(newdata, group_trial == i)
  Results = crqa(currentpair$HR1, currentpair$HR2, delay = optdelay_hr, embed = optemddim_hr, radius = optradius_hr, normalize = 0, rescale = 0, mindiagline = 2, minvertline = 2)
  
  temp_df <- data_frame(group_trial = unique(currentpair$group_trial), study = unique(currentpair$study), group = unique(currentpair$group), trial = unique(currentpair$trial), condition = unique(currentpair$condition), RR = Results$RR, DET = Results$DET, NRLINE = Results$NRLINE, LMAX = Results$maxL, L = Results$L,ENTR = Results$ENTR, rENTR = Results$rENTR, LAM = Results$LAM, TT = Results$TT)
if (i == 1){
    crqa_df_hr <- temp_df
  } else {
    crqa_df_hr <- rbind(crqa_df_hr, temp_df)
  }
}

# Loop for Respiration
for (i in pairs){
  currentpair = filter(newdata, group_trial == i)
  Results = crqa(currentpair$Resp1, currentpair$Resp2, delay=optdelay_resp, embed=optemddim_resp, radius=optradius_resp, normalize=0, rescale=0, mindiagline = 2, minvertline = 2)

  temp_df <- data_frame(group_trial = unique(currentpair$group_trial), study = unique(currentpair$study), group = unique(currentpair$group), trial = unique(currentpair$trial),condition = unique(currentpair$condition), RR = Results$RR, DET = Results$DET, NRLINE = Results$NRLINE, LMAX = Results$maxL, L = Results$L,ENTR = Results$ENTR, rENTR = Results$rENTR, LAM = Results$LAM, TT = Results$TT)
if (i == 1){
    crqa_df_resp <- temp_df
  } else {
    crqa_df_resp <- rbind(crqa_df_resp, temp_df)
  }
}

mean(crqa_df_hr$RR)
mean(crqa_df_resp$RR)
```

Tip: remember to first assess optimal parameters (dimensions, delay, radius) across all timeseries. Tip: it will often fail, just take whatever parameters you get, select optimal across timeseries parameters and run crqa on all timeseries with those. Tip: double check the rr. When I ran the loop, I got very low rr, so I adjusted the radius until the average of rr across all pairs was approx. 4%.

### Creating controls: shuffled controls
 - loop through all pairs and conditions
 - shuffle the timeseries (take a timeseries and rearrange its values in a random order). Tip check the sample() function
 - run crqa and save the output. NB. which delay, embed, radius parameters should you use?
 - statistically compare the crqa indexes in real and shuffled pairs
 
```{r, shuffled controls}

# Loop for HR
crqa_df_hr_shuf <- NULL
crqa_df_resp_shuf <- NULL

set.seed(13)

for (i in pairs){
  currentpair = filter(newdata, group_trial == i)
  Results = crqa(sample(currentpair$HR1), sample(currentpair$HR2), delay = optdelay_hr, embed = optemddim_hr, radius = optradius_hr, normalize = 0, rescale = 0, mindiagline = 2, minvertline = 2)
  
  temp_df <- data_frame(group_trial = unique(currentpair$group_trial), study = unique(currentpair$study), group = unique(currentpair$group), trial = unique(currentpair$trial), condition = unique(currentpair$condition), RR = Results$RR, DET = Results$DET, NRLINE = Results$NRLINE, LMAX = Results$maxL, L = Results$L,ENTR = Results$ENTR, rENTR = Results$rENTR, LAM = Results$LAM, TT = Results$TT)
if (i == 1){
    crqa_df_hr_shuf <- temp_df
  } else {
    crqa_df_hr_shuf <- rbind(crqa_df_hr_shuf, temp_df)
  }
}

set.seed(13)

# Loop for Respiration
for (i in pairs){
  currentpair = filter(newdata, group_trial == i)
  Results = crqa(sample(currentpair$Resp1), sample(currentpair$Resp2), delay=optdelay_resp, embed=optemddim_resp, radius=optradius_resp, normalize=0, rescale=0, mindiagline = 2, minvertline = 2)

  temp_df <- data_frame(group_trial = unique(currentpair$group_trial), study = unique(currentpair$study), group = unique(currentpair$group), trial = unique(currentpair$trial),condition = unique(currentpair$condition), RR = Results$RR, DET = Results$DET, NRLINE = Results$NRLINE, LMAX = Results$maxL, L = Results$L,ENTR = Results$ENTR, rENTR = Results$rENTR, LAM = Results$LAM, TT = Results$TT)
if (i == 1){
    crqa_df_resp_shuf <- temp_df
  } else {
    crqa_df_resp_shuf <- rbind(crqa_df_resp_shuf, temp_df)
  }
}

mean(crqa_df_hr_shuf$RR)
mean(crqa_df_resp_shuf$RR)

```


```{r, t-test of real vs. shuffled}

#t-test for heart rate
t.test(crqa_df_hr$RR, crqa_df_hr_shuf$RR)

#t-test for Respiration
t.test(crqa_df_resp$RR, crqa_df_resp_shuf$RR)

```

### TRICKY! Creating controls: surrogate pair controls
 - Per each real pair, identify at least one surrogate pair (matching one of the participants, with somebody doing the same task, but in a different pair). Tip: Malte will share a method to do this on screen.
 - Run crqa on all the surrogate pairs and save the output. NB. which delay, embed, radius parameters should you use?
 - Test whether crqa shows a difference between real and surrogate pairs


```{r, extractor function}
#defining a function to extract rqa values 
rqa_extractor = function(t1, t2, embed, delay, radius){ #making a function which applied the optimal parameters and then saves the rqa results 
  result = try(crqa(t1, t2, embed = embed, delay = delay, radius = radius, normalize = 0, rescale = 0, mindiagline = 2, minvertline = 2, tw = 0, whiteline = FALSE, recpt = FALSE)) 
  if (length(result) > 1){ 
    results_df = data.frame(RR = result[1], DET = result[2], NRLINE = result[3], maxL = result[4], L = result[5], ENTR = result[6], rENTR = result[7], LAM = result[8], TT = result[9]) 
  } 
  else { results_df = data.frame(RR = NA, DET = NA, NRLINE = NA, maxL = NA, L = NA, ENTR = NA, rENTR = NA, LAM = NA, TT = NA) 
  } 
  return(results_df) 
}

```


```{r, surrogate pairs}
#Surrogate pairs (Kenneths kode)
i = 1 
all_dat = newdata
set.seed(13)

for (g in seq(unique(all_dat$group))){ #loop through all the groups 
  g1 = unique(all_dat$group)[g]
  non_g1 = unique(all_dat$group)[unique(all_dat$group)!= g1] #a list of groups which does not include g1
  g2 = sample(non_g1)[1] #randomly select a group which is in the non_g1 vector
  print(g1)

  for (c in unique(all_dat$condition)){ #looping through conditions 
    temp1 = subset(all_dat, group == g1 & condition == c) #e.g. the first group in condition 'turntaking
    temp2 = subset(all_dat, group == g2 & condition == c) #e.g. the second group in condition 'turntaking
      #doing rqa
    result_Resp = rqa_extractor(t1 = temp1$Resp1, t2 = temp2$Resp2, embed = optemddim_resp, 
                                delay = optdelay_resp, radius = optradius_resp)
    result_HR = rqa_extractor(t1 = temp1$HR1, t2 = temp2$HR2, embed = optemddim_hr, 
                              delay = optdelay_hr, radius = optradius_hr)
    colnames(result_Resp) = paste(colnames(result_Resp), "Resp", sep = ".")
    colnames(result_HR) = paste(colnames(result_HR), "HR", sep = ".")
    
    temp = cbind(result_Resp, result_HR)
    temp$condition = c
    temp$group1 = g1
    temp$group2 = g2
    if (i == 1){ #create df
      surPair_rqa = temp
      i = 2 #if you have already done this then don't do it again
    } else { #append to df
      surPair_rqa = rbind(surPair_rqa, temp)
    }
  print(c)
  }
}


```


```{r}
results_sur <- na.omit(surPair_rqa)

#t-test for heart rate
t.test(crqa_df_hr$RR, results_sur$RR.HR)

#t-test for Respiration
t.test(crqa_df_resp$RR, results_sur$RR.Resp)

```


### Testing effects of conditions
 - make a (probably underpowered) mixed model testing effects of the different conditions on heart rate and respiration coordination
 - N.B: would it make sense to include surrogate pairs? and if so how? what would that tell you?
 

```{r}

model_hr <- lmer(RR ~ condition + (1|group), crqa_df_hr)
summary(model_hr)

model_resp <- lmer(RR ~ condition + (1|group), crqa_df_resp)
summary(model_resp)

model_hr1 <- lmer(L ~ condition + (1|group), crqa_df_hr)
summary(model_hr1)

model_resp1 <- lmer(L ~ condition + (1|group), crqa_df_resp)
summary(model_resp1)

model_hr2 <- lmer(DET ~ condition + (1|group), crqa_df_hr)
summary(model_hr2)

model_drive <- lm(crqa_df_hr$RR ~ crqa_df_resp$RR)
summary(model_drive)

test1 <- rename.vars(crqa_df_hr, "RR", "RR.HR")
test2 <- rename.vars(crqa_df_resp, "RR", "RR.Resp")
test_full <- merge(test1, test2)

ggplot(crqa_df_hr, aes()) + 
  geom_point() + 
  geom_line() +
  geom_text(check_overlap = T, position =  position_jitter(height = 2))+
  ggtitle("Lineplot of semantic distance (trial by trial)")

```


### Effects of respiration coordination on heart rate coordination
 - describe how you would test those.
 - Optional: run the models and report them
 

```{r}

```


------------------------------------------------------------------------------------------------------

These are the questions you need to be able to answer at the end of the assignment (aka that you need to submit as part of the portfolio)

1) How do you preprocess heart rate and respiration data? Describe the process. If any data needs to be excluded, list the excluded data and motivate the exclusion.

2) Do you observe interpersonal coordination in heart rate and respiration? Describe your control baseline, the method used to quantify coordination, and the statistical models used to infer whether coordination was higher than in the baseline. Report the results of the models.

3) Do you observe differences in coordination between conditions? Report the models and results.

4) Is respiration coordination a likely driver of heart rate coordination? Describe how you would test for it. Bonus points if you actually run the tests and report methods and results.

N.B. Because of the timing, we're starting this exercise before collecting the data.
Instead, you will develop your script this week on data from two years ago (Study1) and last year (Study2).
When you hand in the assignment for feedback, you can use the old data. But when you hand in the final version for the exam, you need to adapt your script to use the data we collect next week in the lab.
(For the old data): Note that synchronouns and turn-taking are the same task across both studies, but the third condition is different: two years ago it was self-paced joint reading; last year it was tv-series conversation.

NB: For this exercise, you will need to do something very similiar to what you've done before spread over several weeks. Ie parse data, look at the plots, decide on data cleaning, build a model, and finally evaluate and interpret the results of the models. Going back and copying the approach from previous exercises will likely be a great help.

PROBLEMS BABY:
- "Optimal Radius Not found: try again choosing a wider radius span and larger sample size"
- Loading all data pls

QUESTIONS
- min.rec & max.rec = what is?
- thresholds = er de fornuftige?
- hvilke kan fjernes? kig plots - f.eks. G1, G2? 
- self-paced condition - hvad er den, og skal vi forholde os til alle 4 conditions?
- CRQA: 1 pr. group pr. condition? kan man lave en form for group.by?
    - Plots - hvad viser de?
    - "Images" - er det svarende til dem vi kiggede på på tavlen?
- "explore the lags of coordination" - hvad gør disse fem linjer?
- optimal parameters - gennemsnit af dem vi kan få? median? skal vi helst runde op eller ned?
- statistical comparisons of CRQA - how? t-test? anova? HOW BABY? kigger vi på RR eller L eller alle værdierne?

CODE DUMP:
```{r, CODE DUMP}
g1_self <- read.csv("Study1_G1_T3_SelfPaced.csv")
g1_sync <- read.csv("Study1_G1_T1_Synchronous.csv")
g1_turn <- read.csv("Study1_G1_T2_TurnTaking.csv")

g1_self_1 <- g1_self %>%
  group(n = 100, method = 'greedy') %>% 
  dplyr::summarise(time = mean(time,na.rm=T), 
                   HR1 = mean(HR1,na.rm=T), 
                   HR2 = mean(HR2,na.rm=T), 
                   Resp1 = mean(Resp1,na.rm=T), 
                   Resp2 = mean(Resp2,na.rm=T))

g1_sync_1 <- g1_sync %>%
  group(n = 100, method = 'greedy') %>% 
  dplyr::summarise(time = mean(time,na.rm=T), 
                   HR1 = mean(HR1,na.rm=T), 
                   HR2 = mean(HR2,na.rm=T), 
                   Resp1 = mean(Resp1,na.rm=T), 
                   Resp2 = mean(Resp2,na.rm=T))

g1_turn_1 <- g1_turn %>%
  group(n = 100, method = 'greedy') %>% 
  dplyr::summarise(time = mean(time,na.rm=T), 
                   HR1 = mean(HR1,na.rm=T), 
                   HR2 = mean(HR2,na.rm=T), 
                   Resp1 = mean(Resp1,na.rm=T), 
                   Resp2 = mean(Resp2,na.rm=T))

g1_self_1$HR1 <- scale(g1_self_1$HR1)
g1_self_1$HR2 <- scale(g1_self_1$HR2)
g1_self_1$Resp1 <- scale(g1_self_1$Resp1)
g1_self_1$Resp2 <- scale(g1_self_1$Resp2)

g1_sync_1$HR1 <- scale(g1_sync_1$HR1)
g1_sync_1$HR2 <- scale(g1_sync_1$HR2)
g1_sync_1$Resp1 <- scale(g1_sync_1$Resp1)
g1_sync_1$Resp2 <- scale(g1_sync_1$Resp2)

g1_turn_1$HR1 <- scale(g1_turn_1$HR1)
g1_turn_1$HR2 <- scale(g1_turn_1$HR2)
g1_turn_1$Resp1 <- scale(g1_turn_1$Resp1)
g1_turn_1$Resp2 <- scale(g1_turn_1$Resp2)

g1_self_1$HR1 <- cillitbang(g1_self_1$HR1, threshold_hr)
g1_self_1$HR2 <- cillitbang(g1_self_1$HR2, threshold_hr)

g1_sync_1$HR1 <- cillitbang(g1_sync_1$HR1, threshold_hr)
g1_sync_1$HR2 <- cillitbang(g1_sync_1$HR2, threshold_hr)

g1_turn_1$HR1 <- cillitbang(g1_turn_1$HR1, threshold_hr)
g1_turn_1$HR2 <- cillitbang(g1_turn_1$HR2, threshold_hr)

optimizeParam(na.omit(g1_sync_1$HR1), na.omit(g1_sync_1$HR2), par, min.rec = 3.5, max.rec = 4.5)
  # delay <- 0
  # radius <- 0.2599931
  # emddim <- 20

optimizeParam(na.omit(g1_sync_1$Resp1), na.omit(g1_sync_1$Resp2), par, min.rec = 3.5, max.rec = 4.5)
  # radius <- 0.3481291
  # emddim <- 2
  # delay <- 19

optimizeParam(na.omit(g1_self_1$Resp1), na.omit(g1_self_1$Resp2), par, min.rec = 3.5, max.rec = 4.5)
optimizeParam(na.omit(g1_turn_1$Resp1), na.omit(g1_turn_1$Resp2), par, min.rec = 3.5, max.rec = 4.5) #

optimizeParam(na.omit(g1_self_1$HR1), na.omit(g1_self_1$HR2), par, min.rec = 3.5, max.rec = 4.5) #
optimizeParam(na.omit(g1_turn_1$HR1), na.omit(g1_turn_1$HR2), par, min.rec = 3.5, max.rec = 4.5)

#median(crqa_df_hr$RR)
#median(crqa_df_resp$RR)

all_sync <- newdata %>%
  filter(condition == "Synchronous")

all_conv <- newdata %>%
  filter(condition == "Conversation")

all_turn <- newdata %>%
  filter(condition == "TurnTaking")

all_sync <- unique(all_sync$group_trial)
all_conv <- unique(all_conv$group_trial)
all_turn <- unique(all_turn$group_trial)

all_sync
all_conv
all_turn

lead(unique(all_turn), n = 1L)

for (each in newdata$group_trial){
  
  person_1 <- each
  
  person_2 <- lead(unique(all_turn), n = 1L)
}

or (i in pairs){
  currentpair = filter(newdata, group_trial == i)
  Results = crqa(sample(currentpair$HR1), sample(currentpair$HR2), delay = optdelay_hr, embed = optemddim_hr, radius = optradius_hr, normalize = 0, rescale = 0, mindiagline = 2, minvertline = 2)
  
  temp_df <- data_frame(group_trial = unique(currentpair$group_trial), study = unique(currentpair$study), group = unique(currentpair$group), trial = unique(currentpair$trial), condition = unique(currentpair$condition), RR = Results$RR, DET = Results$DET, NRLINE = Results$NRLINE, LMAX = Results$maxL, L = Results$L,ENTR = Results$ENTR, rENTR = Results$rENTR, LAM = Results$LAM, TT = Results$TT)
if (i == 1){
    crqa_df_hr_shuf <- temp_df
  } else {
    crqa_df_hr_shuf <- rbind(crqa_df_hr_shuf, temp_df)
  }
}
```


